# 1
```{r}
library(tidyverse)
garlic <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/garlic.csv")
glimpse(garlic)
```

# 2

## a

Verifying normality:

```{r}
garlic_lm <- lm(BulbWt ~ factor(Fertilizer), data = garlic)
ggplot(garlic, aes(x = BulbWt, fill = factor(Fertilizer))) +
    geom_density(alpha = 0.2, position = "identity") +
    labs(x = "Bulb Weight")
```

-   Groups do not appear to be Normally distributed

Verifying equal variance:

```{r}
fligner.test(BulbWt ~ factor(Fertilizer), data = garlic)$p.value
```

-   Through the Fligner test, we do not have enough evidence to claim that there are different variances

```{r}
kruskal.test(BulbWt ~ factor(Fertilizer), data = garlic)
```

-   Given the p-value, we believe that there is a significant difference in location. Values in one or more of the groups tend to be higher than others

## b

4 choose 2 so 6 tests

## c

```{r}
1 - (1 - 0.05)^6
```

We solve this problem of experiment-wise error rate by using Tukey's Honest Significant Difference.

## d

```{r}
garlic_aov <- aov(BulbWt ~ factor(Fertilizer), data = garlic)

tukey_garlic <- TukeyHSD(garlic_aov)
print(tukey_garlic)
```

-   Fertilizers 4 and 1 as well as Fertilizers 4 and 3 are statistically different from each other. Pairs 2-1, 3-1, 3-2, 4-2 are not statistically different.

# 3

```{r}
bottle <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/bottle.csv")
glimpse(bottle)
```

Verifying assumption of normality:

```{r}
ggplot(bottle, aes(sample = Units, color = factor(Line))) +
    stat_qq() +
    stat_qq_line()
```

-   The groups are approximately Normal

Verifying assumption of equal variances:

```{r}
library(car)
leveneTest(Units ~ factor(Line), data = bottle)
```

-   There is no evidence to say that the variances between the lines are different

```{r}
bottle_lm <- lm(Units ~ factor(Line), data = bottle)

anova(bottle_lm)
```

-   At least one of the lines has a significant difference in mean units produced

```{r}
TukeyHSD(aov(Units ~ factor(Line), data = bottle))
```

-   We can confirm that 3 seems to be performing worse than 1 or 2 as there are statistical differences between the pairs 

# 4

```{r}
trials <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/trials.csv")
glimpse(trials)
```

```{r}
ggplot(trials, aes(sample = BPChange, color = Treatment)) +
    stat_qq() +
    stat_qq_line()
```

```{r}
leveneTest(BPChange ~ factor(Treatment), data = trials)
```

-   Difference in variances

```{r}
oneway.test(BPChange ~ factor(Treatment), data = trials)
```

-   Difference in means is significant

```{r}
TukeyHSD(aov(BPChange ~ factor(Treatment), data = trials))
```

