```{r}
library(tidyverse)
library(UsingR)
data(normtemp)
glimpse(normtemp)
?normtemp
```

1.  Revisit the NormTemp dataset from Lab 1, where we examined the observed 
mean body temperature (temperature) in comparison to the well-known 
“average” of 98.6


```{r}
t.test(normtemp$temperature, mu = 98.6)
```

-   The p-value is 2.411e-07
-   The p-value is the probability that we observe our data given that the mean body temperature is 98.6
-   Given our low p-value at a significance level of $\alpha = 0.05$ we reject our null hypothesis. We believe the actual mean is significantly different from 98.6
-   The 95% confidence interval for temperature is [98.12200, 98.37646]

```{r}
females <- normtemp[normtemp$gender == 2, ]
t.test(females$temperature, mu = 98.6)
```

-   Given our p-value $< \alpha$ our conclusion does not change

```{r}
ggplot(normtemp, aes(sample = temperature, color = gender)) +
    stat_qq() +
    stat_qq_line()
```

-   Approximately Normal

```{r}
var.test(temperature ~ gender, data = normtemp)
```

- We believe variances are equal

```{r}
t.test(temperature ~ gender, data = normtemp)
```

-   At a significance level of 0.05, we reject our null hypothesis that there is no difference in means between the two genders

1.  The Airline dataset contains information regarding the number of international 
airline travelers (variable air) across different months of the year from 
1949-1960. We are interested in knowing if during this time period there was a
significant difference between air travel in the Summer months of June,
July, and August vs. the remainder of the year? Use a statistical 
hypothesis test (alpha=0.05) to support your answer.

```{r}
data(AirPassengers)
library(tseries)
library(forecast)
cycle(AirPassengers)
```

```{r}
air1 = data.frame(AirPassengers)
air2 = air1 |> mutate(summer = ifelse(cycle(AirPassengers) %in% 6:8, 1, 0))
```

```{r}
ggplot(air2, aes(sample = AirPassengers, color = factor(summer))) +
    stat_qq() +
    stat_qq_line()
```

-   Normality not met. We will use a nonparametric test

```{r}
ggplot(air2, aes(x = AirPassengers, color = factor(summer))) +
    geom_density()
```

-   After plotting the distributions of both groups, we can see a similar shape but not a similar enough variation between the two groups.
-   When we are conducting the Wilcoxon test we can't necessarily claim anything about the mean or median but moreso about the distributional dominance

```{r}
wilcox.test(AirPassengers ~ summer, data = air2)
```

-   At a significance level of 0.05, we reject our null hypothesis that the true location shift is equal to 0