# 1

```{r}
library(AppliedPredictiveModeling)
data(FuelEconomy)
```

```{r}
cor(cars2010[, c("EngDispl", "NumCyl", "ExhaustValvesPerCyl", "VarValveTiming", "FE")])
pairs(cars2010[, c("EngDispl", "NumCyl", "ExhaustValvesPerCyl", "VarValveTiming", "FE")])
engdispl_lm <- lm(FE ~ EngDispl, data = cars2010)
par(nfrow = c(2, 2))
plot(engdispl_lm)
summary(engdispl_lm)
```

-   Highest correlation with `FE` is `EnglDispl`

```{r}
cor.test(cars2010$FE, cars2010$EngDispl)
```

-   P-value is significant at a 0.05 level and we have evidence that the correlation coefficient is not equal to 0

```{r}
cor(cars2010[, c("EngDispl", "NumCyl", "ExhaustValvesPerCyl", "VarValveTiming")])
```

-   `EnglDispl` and `NumCyl` have a large correlation between one another

```{r}
slr <- lm(FE ~ EngDispl, cars2010)
summary(slr)$r.squared
```

-   F-statistic is 1803 with a p-value of 2.2e-16. Overall model is significant
-   $y = 50.5632 - 4.5209x_1$ 
-   $R^2$ is 0.620 which means that 62% of the variability in `FE` can be explained by `EngDispl` alone

# 2

```{r}
library(tidyverse)
library(ggplot2)
icecream <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/icecream.csv", sep = " ")
glimpse(icecream)
```

```{r}
sales_slr <- lm(Sales ~ Temperature, icecream)
par(mfrow = c(2, 2))
plot(sales_slr)

ggplot(icecream, aes(x = Temperature, y = Sales)) +
    geom_point() +
    stat_smooth(method = "lm")

summary(sales_slr)
```

-   Residuals seem to be Normally distributed from the QQ-Plot
-   Error variance is constant and does not show a pattern
-   Linearity of the mean
-   Overall F-test is significant so `Temperature` seems to be significant in explaining `Sales`
-   Parameter estimate for `Temperature` is 1.0889

# 3

```{r}
minntemp <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/minntemp.csv", sep = " ")
glimpse(minntemp)
```

```{r}
temp_slr <- lm(Temp ~ Time, minntemp)
par(mfrow = c(2, 2))
plot(temp_slr)
ggplot(minntemp, aes(x = Time, y = Temp)) +
    geom_point() +
    stat_smooth(method = "lm")
summary(temp_slr)
```

-   No linearity of the mean, seems quadratic
-   No constant variance
-   Normally distributed errors
-   No statistical evidence that `time` is related to `temperature` at a confidence level of 0.05