---
title: Getting to Know Your Classmates Data
author: Carlos Castillo, Yang Chen, Nick Hicks, Manisha Kasam
date: June 30, 2023
---

# Preparing and Loading Data
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)
library(countrycode)
library(ggthemes)

# setwd("summer_r")
classmates <- read.csv("data/Get_to_know_survey_2023.csv")
glimpse(classmates)
```

# Birth_Month Column
​
```{r}
classmates <- classmates %>%
    rowwise() %>%
    mutate(Birth_Month = str_split(Birth_Month_Year, ", ")[[1]][1])

classmates %>%
    select(Birth_Month)
```

# Birth_Year Column
​
```{r}
classmates <- classmates %>%
    rowwise() %>%
    mutate(Birth_Year = str_split(Birth_Month_Year, ", ")[[1]][2])

classmates %>%
    select(Birth_Year)
```

# Standardizing Birth_Year and Birth_Month
​
```{r}
classmates <- classmates %>%
    mutate(Birth_Month = str_to_title(str_trim(Birth_Month))) %>%
    mutate(Birth_Month = ifelse(Birth_Month %in% as.character(1:12), month.name[as.integer(Birth_Month)], Birth_Month))

classmates <- classmates %>%
    mutate_at(c("Birth_Year"), ~ na_if(., ""))

classmates %>%
    select(Birth_Month, Birth_Year)
```

# 5 Most Frequent Birth Months, 5 Least Frequent
​
```{r}
classmates %>%
    filter(!is.na(Birth_Month)) %>%
    count(Birth_Month) %>%
    arrange(desc(n)) %>%
    head(n = 5)

classmates %>%
    filter(!is.na(Birth_Month)) %>%
    count(Birth_Month) %>%
    arrange(n) %>%
    head(n = 5)
```

# 5 Most Frequent Birth Years, 5 Least Frequent

```{r}
classmates %>%
    count(Birth_Year) %>%
    arrange(desc(n)) %>%
    head(n = 5)

classmates %>%
    count(Birth_Year) %>%
    arrange(n) %>%
    head(n = 5)
```

# Standardizing Country Column

We are standardizing `Country` first. We decided that we would also be filtering out any US state with a non-US country combination.

```{r}
classmates <- classmates %>%
    mutate_at(c("Country"), ~ na_if(., "")) %>%
    mutate(Country = str_to_title(str_trim(Country))) %>%
    mutate(Country = str_replace_all(Country, c("[Uu]\\.*s\\.*a*\\.*" = "United States", ".*United States*.*" = "United States"))) %>%
    mutate(Country = ifelse(is.na(Country), "United States", Country)) %>%
    mutate(Country = countrycode(Country, "country.name", "country.name"))

classmates %>%
    filter(!is.na(Country)) %>%
    count(Country) %>%
    arrange(desc(n)) %>%
    head(n = 5)

classmates %>%
    filter(!is.na(Country)) %>%
    count(Country) %>%
    arrange(n) %>%
    head(n = 5)
```

# Filtering Out Non-US States, Standardizing States

```{r}
classmates <- classmates %>%
    mutate(State = str_trim(State)) %>%
    mutate(State = case_when(
        str_detect(Country, "United States", negate = TRUE) ~ NA_character_,
        (str_to_upper(State) %in% state.abb) ~ state.name[match(State, state.abb)],
        (str_to_title(State) %in% state.name) ~ State,
        TRUE ~ NA_character_
    )) %>%
    mutate(State = str_to_title(State))

classmates %>%
    select(State)
```

# Five Most Frequent States, Five Least Frequent States

```{r}
classmates %>%
    filter(!is.na(State)) %>%
    count(State) %>%
    arrange(desc(n)) %>%
    head(n = 5)

classmates %>%
    filter(!is.na(State)) %>%
    count(State) %>%
    arrange(n) %>%
    head(n = 5)
```

# Cleaning `How_Hear_About_Program` Column

```{r}
classmates <- classmates %>%
    mutate(How_Hear_About_Program = str_trim(How_Hear_About_Program)) %>%
    mutate(How_Hear_About_Program = paste(str_split(How_Hear_About_Program, " ")[[1]][-1], collapse = " "))

classmates %>%
    filter(!is.na(How_Hear_About_Program)) %>%
    count(How_Hear_About_Program) %>%
    arrange(desc(n)) %>%
    head(n = 5)
```

# Wide to Long `Languages` Column

```{r}
classmates_copy <- data.frame(classmates)
classmates <- classmates %>%
    mutate(Languages = str_squish(Languages)) %>%
    separate_longer_delim(Languages, ", ") %>%
    mutate(Languages = str_replace_all(Languages, ",", ""))

classmates

languages_top_5 <- classmates %>%
    filter(!is.na(Languages)) %>%
    count(Languages) %>%
    arrange(desc(n)) %>%
    head(5)

languages_top_5
```

# Extra Cleaning & Analyses

## Top 5 Languages Visualized

```{r}
ggplot(languages_top_5, aes(x = reorder(Languages, desc(n)), y = n, fill = Languages)) +
    geom_col() +
    labs(x = "Languages", y = "Counts", title = "MSA Class of 2024: Top 5 Languages") +
    theme_economist()
```

## Wide to Long `Hobbies` Column

```{r}
classmates <- classmates %>%
    mutate(Hobbies = str_squish(Hobbies)) %>%
    separate_longer_delim(Hobbies, ", ")

classmates

classmates_copy <- classmates_copy %>%
    mutate(Hobbies = str_squish(Hobbies)) %>%
    separate_longer_delim(Hobbies, ", ")

hobbies_top_5 <- classmates_copy %>%
    filter(!is.na(Hobbies)) %>%
    count(Hobbies) %>%
    arrange(desc(n)) %>%
    head(n = 5)

hobbies_top_5
```

## Top 5 Hobbies Visualized

Note: Using copy of older `classmates` data frame since we don't want to "double-count" student hobbies after making the `language` column long. Maybe there's a better way to do this through grouping.

```{r}
ggplot(hobbies_top_5, aes(x = reorder(Hobbies, desc(n)), y = n, fill = Hobbies)) +
    geom_col() +
    labs(x = "Hobbies", y = "Counts", title = "MSA Class of 2024: Top 5 Hobbies") +
    theme_economist()
```

## Standardizing Birth_Month_Year
```{r}
classmates <- classmates %>%
    mutate(Birth_Month_Year = paste(Birth_Month, Birth_Year, sep = ", "))

classmates %>%
    select(Birth_Month_Year)
```

## Student Countries Heatmap

```{r warning=FALSE, message=FALSE}
library(rnaturalearth)
library(rnaturalearthdata)

selected_countries <- classmates %>%
    select(Country) %>%
    distinct(Country) %>%
    arrange(Country)
world_data <- ne_countries(scale = "medium", returnclass = "sf")
selected_countries_data <- world_data %>%
    inner_join(selected_countries, by = c("name" = "Country"))

ggplot() +
    geom_sf(data = world_data, color = "grey50", fill = "lightblue") +
    geom_sf(data = selected_countries_data, fill = "red", alpha = 0.7) +
    labs(
        title = "Countries of Origin",
        subtitle = "MSA 2024",
        caption = "Library: Natural Earth"
    ) +
    theme_minimal()
```