---
title: Lab 6
date: 07/06/2023
---

# 1

```{r}
library(tidyverse)
library(AppliedPredictiveModeling)
data(FuelEconomy)
glimpse(cars2010)
```

## a

```{r}
cars2010 <- cars2010 %>%
    mutate(across(!c(EngDispl, FE), as.factor))
cars_lm <- lm(FE ~ ., data = cars2010)
summary(cars_lm)
```

-   The F p-value is significant meaning that our overall model is significant in predicting `FE`
-   The 13 variables explain 83.33 percent of variation in fuel economy

## b

```{r}
car::Anova(cars_lm)
```

-   Highest p-value is `VarValveTiming` with 0.584

## c

```{r}
cars_2010_sub <- cars2010 %>%
    select(-VarValveTiming)

cars_lm2 <- lm(FE ~ ., cars_2010_sub)
summary(cars_lm2)
car::Anova(cars_lm2)
```

-   The p-value for the model did not change significantly
-   $R^2$ did not change significantly and $R_a^2$ increased a little 
-   

## d

Dropping `IntakeValvePerCyl`:

```{r}
cars_2010_sub <- cars_2010_sub %>%
    select(-IntakeValvePerCyl)

cars_lm3 <- lm(FE ~ ., cars_2010_sub)
summary(cars_lm3)
car::Anova(cars_lm3)
```

```{r}
cars_2010_sub <- cars_2010_sub %>%
    select(-TransLockup)

cars_lm4 <- lm(FE ~ ., cars_2010_sub)
summary(cars_lm4)
car::Anova(cars_lm4)
```

```{r}
cars_2010_sub <- cars_2010_sub %>%
    select(-VarValveLift)

cars_lm5 <- lm(FE ~ ., cars_2010_sub)
summary(cars_lm5)
car::Anova(cars_lm5)
```

-  Both $R^2$ and $R_a^2$ decreased
-  9 variables significant at the 0.008 level 


