---
title: Lab 12
date: 07/17/2023
---

```{r}
library(tidyverse)
safety <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/safety.csv")
```

# 1

## a

-   `Unsafe` is ordinal
-   `Type` is nominal
-   `Region` is ordinal
-   `Weight` is ordinal
-   `Size` is ordinal

## b

```{r}
table(safety$Region, safety$Unsafe)

ggplot(safety) +
    geom_bar(aes(x = Region, fill = factor(Unsafe)))
```

-   Percentage of cars classified as unsafe in the Asia region is 42.9 percent
-   Percentage of cars classified as safe in the North America region is 75.4 percent
-   Appropriate test to use is Mantel-Haenszel
    -   $H_0:$ No linear association, $H_a:$ Linear association exists

```{r}
library(vcdExtra)
library(DescTools)

CMHtest(table(safety$Region, safety$Unsafe))$table[1, ]

OddsRatio(table(safety$Region, safety$Unsafe))

table(safety$Region, safety$Unsafe)
```

-   At an $\alpha$ level of 0.05, we do not reject the null hypothesis that there is no linear association between `Region` and `Safety`
-   There are 0.435 times the odds to be safe if it's in the Asia region

## c

Appropriate test is Mantel-Haenszel

```{r}
CMHtest(table(safety$Size, safety$Unsafe))$table[1, ]
```

-   We reject the null hypothesis that there is no linear association between the two variables

```{r}
cor.test(
    x = as.numeric(ordered(safety$Size)), 
    y = as.numeric(ordered(safety$Unsafe)), 
    method = "spearman"
)
```

-   It's not too strong at $\rho = -0.542$